Making Sequential Programs Deterministic

### 2019-03-27 design.tex
moving text around
precision

\begin{itemize}
    \item Problem with \<Map.get()> : Annotating as \<@PolyDet> is imprecise.
    \item We circumvent this by hardcoding rules to make it precise. (Describe the rule here)
\end{itemize}

\subsection{Determinism checker specific rules for precision}\label{precision}

\begin{itemize}
    \item \<Set.equals()> is annotated as \<@Poly("up")>. This is sound but imprecise.
    \item Example1: \codeid{@OrderNonDet Set<@Det List>.equals(@OrderNonDet Set<@Det List>)} is deterministic
    \item Example2: \codeid{@OrderNonDet Set<@OrderNonDet List>.equals(@OrderNonDet Set<@OrderNonDet List>)} is not deterministic.
    \item Explain the rules under which we make Set equality deterministic.
    \item We annotate System.getProperty as \<@NonDet> because the results could vary across machines.
    \item Exceptions: line.Sepaator, file.Separator. and path.Separator.
    \item discuss type on Exceptions and unsoundness due to this.
\end{itemize}

>>>>

\subsection{Determinism checker specific rules for precision}\label{precision}
We annotated the \codieid{equals()} method of the \<Set> class in the JDK as
\begin{verbatim}
PolyDet$\uparrow$ boolean equals(PolyDet Set<E> this, PolyDet Object o)
\end{verbatim}
While this is sound, it is imprecise in the case of \codeid{OrderNonDet Sets}. We added the following rule shown in Figure~\ref{fig-precision-rules} to our checker to improve precision:

\begin{figure}
    \todo{linebreak}
    $\infer{\Gamma \vdash \mathit{z : Det}}{\Gamma \vdash \mathit{x} : \mathit{OrderNonDet} \ \mathit{Set}\langle \kappa_{x} \ \beta_{x}\rangle, \mathit{y} : \mathit{OrderNonDet} \ \mathit{Set}\langle \kappa_{y} \ \beta_{y}\rangle & \mathit{z = x.equals(y)} & \langle \kappa_{x} \ \beta_{x}\rangle \nsqsubseteq \langle \mathit{OrderNonDet \ List} \rangle & \langle \kappa_{y} \ \beta_{y}\rangle \nsqsubseteq \langle \mathit{OrderNonDet \ List} \rangle }$

    \caption{Improving precision}
    \label{fig-precision-rules}
\end{figure}
According to this rule, if the \<Set> receiver has type \codeid{@OrderNonDet Set<@Det List} and the argument to the \codeid{equals()} method is
also of the same type, the return type gets the qualifier \<Det>. Without this precision rule, the return type would have resolved to \<NonDet> which
is sound but imprecise. This rule is careful enough to not annotate the return type of \<equals()> as \<Det> if a type argument to either the
receiver or the argument happens to be an \codeid{OrderNonDet List} (An order nondeterminstic list does not provide any guarantee on the
equality as it is dependent on the iteration order).

\begin{itemize}
    \item Problem with \<Map.get()> : Annotating as \<@PolyDet> is imprecise.
    \item We circumvent this by hardcoding rules to make it precise. (Describe the rule here)

    \item We annotate System.getProperty as \<@NonDet> because the results could vary across machines.
    \item Exceptions: line.Separator, file.Separator. and path.Separator.
    \item discuss type on Exceptions and unsoundness due to this.
\end{itemize}
